---
title: "labclassweek08"
author: "Stefanio Marques - 25250375"
format: pdf
editor: visual
---

### Lab 7 (Topic 6) - Handling different data types

#### Introduction

```{r}
library(MASS)
library(tidyverse)
library(dplyr)

data(quine)

# Attach the data set to the R search path
attach(quine)

# Look at the structure of the data
str(quine)
```

#### Question 02
Store the data into a tibble and give the columns more informative names:

```{r}

# Convert the data frame to a tibble and rename columns
quine_tbl <- as_tibble(quine) |>
  rename(
    Ethnicity = Eth,
    Gender = Sex,
    AgeGroup = Age,
    LearnerStatus = Lrn,
    DaysAbsent = Days
  )

# View the new tibble
print(quine_tbl)
```

#### Question 03
Use the table function on the ethnic background variable. What ethnicity is most prevalent in the data set?

```{r}
table(quine_tbl$Ethnicity)
```

#### Question 04
Use the factor function to change the age and learner variables into (sensibly!) ordered factors using the extra argument ordered = TRUE. Check the structure of the data set to ensure what you think has happened to these variables, actually has.

```{r}
# The levels F0, F1, F2, F3 represent age classes in increasing order.
quine_tbl$AgeGroup <- factor(quine_tbl$AgeGroup,
                             levels = c("F0", "F1", "F2", "F3"),
                             ordered = TRUE)

# Change LearnerStatus to an ordered factor
# We'll order it as "Slow Learner" < "Average Learner"
quine_tbl$LearnerStatus <- factor(quine_tbl$LearnerStatus,
                                  levels = c("SL", "AL"),
                                  ordered = TRUE)

# Check the structure to see the changes
str(quine_tbl)
```


#### Question 05
Use the table() to compare learner ability and ethnicity in a 2-way table. Now use dplyr::count() to get the same information.

```{r}
# 1. Using table()
table(quine_tbl$LearnerStatus, quine_tbl$Ethnicity)

# 2. Using dplyr::count()
quine_tbl |>
  count(LearnerStatus, Ethnicity)
```

#### Question 06
Use aggregate() to calculate the mean days absent by learner status and sex. Now use functions from dplyr to get the same information.

```{r}
# 1. Using aggregate()
aggregate(DaysAbsent ~ LearnerStatus + Gender, data = quine_tbl, FUN = mean)

# 2. Using dplyr
quine_tbl |>
  group_by(LearnerStatus, Gender) |>
  summarise(MeanDaysAbsent = mean(DaysAbsent, na.rm = TRUE))
```

#### Question 07
Use the cut() function on the days absent variable to divide it up into 3 categories: few (0-20), medium (20-60) and many (60-100). Have a look at the output. Is there anything strange going on here? If so, how can you fix it (use the help file!). How many students fall into the many category?

```{r}
# First, let's check the range of DaysAbsent
summary(quine_tbl$DaysAbsent)
# Min = 0, Max = 81. The 60-100 range is fine.

# Use cut()
days_cats <- cut(quine_tbl$DaysAbsent,
                 breaks = c(0, 20, 60, 100),
                 labels = c("few", "medium", "many"))

# Look at the output
table(days_cats)
```

Is there anything strange going on here?

Yes. If you look at the summary() of DaysAbsent, the minimum value is 0. By default, cut() creates intervals that are open on the left ((a, b]). This means the interval (0, 20] would exclude any student who had 0 days absent.

How can you fix it?

You can fix this by using the include.lowest = TRUE argument. This ensures the first bin is "closed" on the left, making it 0, 20.

```{r}
# Corrected cut() function
days_cats_fixed <- cut(quine_tbl$DaysAbsent,
                       breaks = c(0, 20, 60, 100),
                       labels = c("few", "medium", "many"),
                       include.lowest = TRUE) # This is the fix

# Look at the corrected table
table(days_cats_fixed)
```

#### Question 08
Load the gapminder dataset from the gapminder package. Focus on the 2007 data and European counties. Do a plot with life expectancy on the x-axis, and the countries sorted by according their life expectancy values on the y-axis.

```{r}
library(gapminder)
library(ggplot2)
library(dplyr) # Should already be loaded

# Filter the data for European countries in 2007
gapminder_2007_europe <- gapminder %>%
  filter(year == 2007, continent == "Europe")

# Create the plot
ggplot(gapminder_2007_europe, aes(x = lifeExp, y = reorder(country, lifeExp))) +
  geom_point() +
  labs(
    title = "Life Expectancy in Europe (2007)",
    x = "Life Expectancy (Years)",
    y = "Country"
  ) +
  theme_minimal() # A clean theme
```

#### Question 09

Based on an analysis of the R-help mailing list archive for May 2020:

Topics: There were 5 unique topics (threads) discussing the new version.

3 topics matched "R 4" (e.g., "Partial matching list elements in R 4.0")
1 topic matched "R-4" (e.g., "PCRE configure problem with R-4.0.0")
1 topic matched "R/4" (e.g., "Problem with checks on R/4.0.0")

Messages: There were a total of 21 messages posted across these 5 topics.

11 messages for "R 4" topics
7 messages for the "R-4" topic
3 messages for the "R/4" topic
