---
title: "Worksheet1"
author: "Stefanio Marques"
date: "'r Sys.Date()'"
format: docx
editor: visual
---

## Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
```

```{r}
setwd("/Users/diegomandu/Downloads/class/programming with R/econometrics labs")
```

```{r}
library(readxl)
library(tidyverse)
library(stargazer)
library(modelsummary)
library(ggplot2)
```

## Question 01

Upload the excel dataset using the code:

```{r}
WAGE1 <- read_xlsx("WAGE1.xlsx")
```

Familiarize yourself with the dataset looking at all the variables

```{r results='hide'}
glimpse(WAGE1)
```

Obtain a table of summary statistics

```{r}
datasummary_skim(WAGE1, output = "default", histogram=T)

caption <- 'Table 1: Descriptive statistics of our sample'
reference <- 'Source: Data are from Wooldridge.'
datasummary_balance(~1,
                    data = WAGE1,
                    title = caption,
                    notes = reference)
```

Produce a boxplot for variables wage, educ, exper, tenure, numdep

```{r}
var1 <- WAGE1$wage
var2 <- WAGE1$educ
var3 <- WAGE1$exper
var4 <- WAGE1$tenure
var5 <- WAGE1$numdep
boxplot(var1,var2,var3,var4,var5)
```

\newpage

## Question 5-6

```{r}
WAGE1 %>%
  group_by(female) %>%
  summarise("Wage by sex"= mean(wage),
            "Education"= mean(educ),
            "Experience"= mean(exper),
            "Tenure"= mean(tenure),
            "Number dependents"= mean(numdep))
```

Boxplot of wage by gender

```{r}
boxplot(wage~female,
        data=WAGE1,
        main="Wage boxplots by gender",
        xlab = "Gender",
        ylab = "Wage",
        col="orange",
        border="brown")
```

To produce a nice-looking table of avarages by group

```{r}
caption2 <- 'Table 2:Descriptive statistics by gender'
reference2 <- 'Source: Data are from Wooldridge.'

datasummary_balance(~female,
                    data = WAGE1,
                    title = caption2,
                    notes = reference2, dinm = FALSE)
```

## Question 7-8

```{r}
p <- ggplot(WAGE1,aes(x=wage)) + geom_density() + geom_vline(aes(xintercept=mean(wage)),
color="blue", linetype="dashed", linewidth=1)
p
```
___

```{r}
WAGE1$lnwage <- log(WAGE1$wage)
boxplot(WAGE1$lnwage)
```

## Question 9-13

```{r}
data <- WAGE1 %>%
  select(lnwage, educ, exper, tenure, numdep)
datasummary_correlation(data)
```

```{r}
ggplot(WAGE1,aes(x=educ,y=lnwage)) +
  geom_point()+
  geom_smooth(se=F)+
  geom_smooth(method="lm",se=F)+
  labs(title = "Relationship between wages and education",
                     subtitle = "Source:own data")
```

regressors

```{r}

```

_____
OLS
_____

```{r}
reg1 <- lm(lnwage ~ educ + tenure + nonwhite + married + exper + northcen +  female + numdep, data=WAGE1)

modelsummary(list(reg1))

reg5 <- lm(lnwage ~ educ_level + tenure + nonwhite + married + exper + northcen +  female + numdep, data=WAGE1)

modelsummary(list(reg5))
```

____

```{r}
library(stargazer)
stargazer(reg1, reg5,
          title = "Results",
          type = "text",
          out = "table1.text",
          header = FALSE,
          no.space = TRUE)
```





