---
title: "Tutorial 6"
author: "Stefanio Marques - 25250375"
date: "`r Sys.Date()`"
format: pdf
editor: visual
---

### Introduction

```{r}
setwd("/Users/diegomandu/Downloads/class/programming with R/econometrics labs/tutorial 6")
```

#### Load the packages

```{r}
library(readxl)
library(tidyverse)
library(modelsummary)
library(ggfortify)
library(car)
library(stargazer)
library(lmtest)
library(estimatr)
library(sandwich)
```

####
Upload the dataset

```{r}
airquality <- read_xlsx("airquality.xlsx")

names(airquality)
```

### Part A - Estimate a linear regression model that explains airq from the other variables using OLS.


```{r}
model <- lm(airq ~ vala + rain + coas + dens + medi, data = airquality)
summary(model)
```

### Part B - Test hypothesis (average income)
Test the null hypothesis that average income does not affect the airquality. Test the
joint hypothesis that none of the variables has an effect upon airquality

```{r}
# To test this hypothesis formally (H0: medi = 0):
linearHypothesis(model, "medi = 0")
```

### Part C - Test joint hypothesis (F-test)
Use the command autoplot to graphically inspect the residuals; is there any sign of
heteroskedasticity?

```{r}
#To test H0 (all coefficients are zero) formally:
linearHypothesis(model, c("vala = 0", "rain = 0", "coas = 0", "dens = 0", "medi = 0"))
```

Graphically inspect residuals

```{r}
autoplot(model) +
  theme_bw(base_size = 14)
```

### Part D - Breusch-Pagan test
Perform a Breusch-Pagan test for heteroskedasticity related to all five explanatory
variables.

```{r}
bptest(model)


bptest(model, ~ vala + rain, data = airquality)
```

### Part E - White test
Perform a White test for heteroskedasticity. How reliable is this test given that we
have 30 observations and how many degrees of freedom on the chi-square
distribution?????

```{r}
bptest(model, ~ vala + rain + coas + dens + medi + I(vala^2) + I(rain^2) + I(dens^2) + I(medi^2), data = airquality)
```

### Part F - Check heteroskedasticity by 'coas'
Check if heteroskedasticity may depend on whether smsa is on the coas:

```{r}
plot(airquality$coas, residuals(model),
     xlab = "Coast (1=Yes, 0=No)", ylab = "Residuals")

# Find percentage of observations with coas=0
(percentage_coas_0 <- mean(airquality$coas == 0))

summary(airquality$coas)

# Perform Goldfeld-Quandt test
model2 <- lm(airq ~ vala + rain + dens + medi, data = airquality)
gq_test <- gqtest(model2, order.by = ~ coas, point = percentage_coas_0, data = airquality)
print(gq_test)

# suppose the order by variable was continuous, for example medi
gq_test2 <- gqtest(model, order.by = ~ medi, point = 0.5, fraction = 0.2, data = airquality)
print(gq_test2)
```

### Part G - Multiplicative Heteroskedasticity
Both BP and White test have quite a generic alternative hypothesis and have
therefore a relatively ow power. Assume now that heteroskedasticity is
multiplicative

```{r}
plot(airquality$vala, residuals(model), xlab = "Value Added", ylab = "Residuals")
plot(airquality$rain, residuals(model), xlab = "Rain", ylab = "Residuals")
plot(airquality$dens, residuals(model), xlab = "Density", ylab = "Residuals")
plot(airquality$medi, residuals(model), xlab = "Median Income", ylab = "Residuals")
```

### Part H - Test for multiplicative heteroskedasticity

```{r}
res_sq <- residuals(model)^2


aux_variance_model <- lm(log(res_sq) ~ coas + medi, data = airquality)
summary(aux_variance_model)

# This is the F-test of the auxiliary regression.
# We test H0: coas = 0 AND medi = 0
linearHypothesis(aux_variance_model, c("coas = 0", "medi = 0"))
```

### Part I - Robust Standard Errors
Obtain the robust standard errors. You need to install estimatr and car

```{r}
model_robust <- lm_robust(airq ~ vala + rain + coas + dens + medi,
                          data = airquality)

modelsummary(list(model,model_robust), stars=TRUE)
```

### Part J - Bootstrapped Standard Errors
Obtain the bootstrapped s.e. Here you need the package sandwich

```{r}
# Uses lmtest and sandwich packages
coeftest(model, vcov = vcovBS(model), R=1000)
```

### Part K - Compute FGLS Estimator
Based on the test in h. compute the FGLS estimator for the linear model. Compare
the results with the OLS estimates

```{r}
variance <- exp(fitted(aux_variance_model))

# Estimate the model using lm and specify weights
model_fgls <- lm(airq ~ vala + rain + coas + dens + medi,
                 weights = 1 / variance,
                 data = airquality)

summary(model_fgls)
```

##### Compare OLS, Robust, and FGLS estimates

```{r}
stargazer(model, model_robust, model_fgls,
           type = "text",
           no.space = TRUE)


# Alternatively

models_list <- list(
  "OLS" = model,
  "Robust (estimatr)" = model_robust,
  "FGLS" = model_fgls
)

modelsummary(models_list, stars = TRUE)
```


