---
title: "Worksheet2"
author: "Stefanio Marques"
date: "`r Sys.Date()`"
format: pdf
editor: visual
---

## Introduction

Clean the Envirioment

```{r}
rm(list = ls())
```

```{r}
setwd("/Users/diegomandu/Downloads/class/programming with R/econometrics labs/tutorial 2")
```

## Load the Packages

```{r}
library(readxl)
library(tidyverse)
library(stargazer)
library(modelsummary)
library(ggplot2)
library(marginaleffects)
library(car)
```

Upload the excel dataset using the code:

```{r}
WAGE1 <- read_xlsx("WAGE1.xlsx")
```

## Question 1-2

Generate the natural log of wages, and run the basic regression we did in previous tutorial:

```{r}
WAGE1$lnwage <- log(WAGE1$wage)
```

```{r}
reg1 <- lm(lnwage ~ educ + exper + tenure + female + nonwhite + married + numdep + northcen + south + west, data = WAGE1)
```

Relationship between two variables

```{r}
ggplot(WAGE1, aes(x=exper,y=lnwage))+
geom_point()+
geom_smooth(se=F)+
geom_smooth(method="lm",se=F)+
labs(
title="Relationship between wages and exper",
subtitle="Source: own data"
)
```

Estimate the new model with a quadratic term for experience

```{r}
reg2 <- lm(lnwage ~ educ + exper + I(exper^2) + tenure + female + nonwhite + married + numdep + northcen + south + west, data = WAGE1)
```

Basic summary of the new model

```{r}
modelsummary(reg2)
```

```{r}
avg_slopes(reg2) #avarage marginal effect
```

```{r}
avg_slopes(reg2, newdata = "mean")
```

## Question 03

Estimating the Interaction Model

```{r}
reg3 <- lm(lnwage ~ educ * female + exper + I(exper^2) + tenure + nonwhite + 
                    married + numdep + northcen + south + west, data = WAGE1)
modelsummary(reg3)
```

\newpage

## Question 04

```{r}
reg4 <- lm(lnwage ~ educ * female + exper + I(exper^2) + tenure + I(tenure^2) + 
                    nonwhite + married + numdep + northcen + south + west, 
           data = WAGE1)
modelsummary(list(reg1, reg2, reg3, reg4))
```

## Question 05

```{r}
library(modelsummary)
modelplot(list("Final Model" = reg4), coef_omit = 'Intercept')
```

Use these commands to plot the marginal effects of education by gender

```{r}
plot_slopes(reg4,
            variables = "educ",
            slope = "dydx",
            by=c("female"))
```

Marginal effects of experience from the specification in 4

```{r}
plot_slopes(reg4,
            variables = "exper",
            condition = c("exper"))
```

Calculate the average return to education across your entire sample

```{r}
avg_slopes(reg4, variables = "educ")
```

## Question 06

```{r}
reg5 <- lm(lnwage ~ educ*female + exper*female + I(exper^2) + tenure + I(tenure^2) + nonwhite + married + numdep + northcen + south + west,
           data = WAGE1)
```

```{r}
modelsummary(list(reg1, reg2, reg3, reg4, reg5))
```

Plot the marginal effect of experience, conditioned on experience level

```{r}
plot_slopes(reg5,
            variables = "exper",
            condition = c("exper", "female"))
```

## Question 07

```{r}
stargazer(reg1, reg5,
          title="Wage Regression Results",
          type="text",
          out="table1.text",
          header=FALSE,
          no.space=TRUE)
```

Alternatively, you can change to hyml file

```{r}
stargazer(reg1, reg5,
          title="Wage Regression Results",
          type="html",
          out="table1.html")
```

## Question 08

- Go to menu File and select Knit document.
- If execution is halted, there is a mistake in the code, go back and try to fix it

