---
title: "Tutorial 4"
author: "Stefanio Marques - 25250375"
date: "`r Sys.Date()`"
format: pdf
editor: visual
---

# Introduction

```{r}
setwd("/Users/diegomandu/Downloads/class/programming with R/econometrics labs/tutorial 4")
```

### Load the packages

```{r}
library(readxl)
library(tidyverse)
library(ggfortify)
library(kdensity)
library(tseries)
library(stargazer)
library(lmtest)
library(modelsummary)
library(ggplot2)
library(marginaleffects)
library(car)
```

### Part 01 (Upload the dataset)

```{r}
WAGE1 <- read_xlsx("WAGE1.xlsx")
```

#### b. Generate the natural log of wages and estimate a model that controls for education, tenure and its square, experience and its square, female, married, numdep and residence dummies.

```{r}
reg1 <- lm(log(wage) ~ educ + tenure + I(tenure^2) + exper + I(exper^2) + 
               female + married + numdep + south + northcen + west, 
               data = WAGE1)

summary(reg1)
```

### Questions 1 - 6 focus on analysis of residuals:

#### Generate a plot of residuals vs fitted values.

```{r}
autoplot(reg1) +
  theme_bw(base_size = 14)

plot(reg1)
```

-   Residuals vs Fitted: The blue line is relatively flat and centered around zero, which is a good sign for the linearity assumption. However, the vertical spread of the points increases as the fitted values get larger. This shape is a sign of heteroscedasticity.

-   Normal Q-Q: The points fall along the dashed diagonal line. This indicates that the residuals follow a normal distribution, satisfying the normality assumption.

-   Scale-Location: This plot also checks for constant variance (homoscedasticity). The upward-sloping blue line confirms what we saw in the first plot: the variance of the residuals increases with the fitted values.

-   Residuals vs Leverage: This plot helps identify influential outliers. There are no points in the upper or lower right corners. This is a good sign, suggesting that no single observation is disproportionately affecting the regression results.

The model shows signs of heteroscedasticity.

The solution might be compute heteroscedasticity-robust standard errors to account for the non-constant variance, making your hypothesis tests valid again.

#### Check for outliers looking at the standardized residuals vs leverage plot. What do you think?

```{r}
testoutlier <- outlierTest(reg1)

print(testoutlier)
```


```{r}
reg2 <- lm(log(wage) ~ educ + I(educ^2) + tenure + I(tenure^2) + exper + I(exper^2) + 
               female + married + numdep + south + northcen + west, 
               data = WAGE1)

summary(reg2)
```

#### Question 02

#### 03 Graph the following:

```{r}
influenceIndexPlot(reg2, vars = "hat", 
                   id=F, main = "Leverage (hat values)")

influenceIndexPlot(reg2, vars = "Studentized",
                   id=F, main = "Studentized Residuals")

influenceIndexPlot(reg2, vars = "Cook",
                   id=F, main = "Cook's distance")
```

Leverage: Observation #265 stands out for having unusual characteristics (high leverage).

Outliers (Studentized Residuals): Observation #128 is the main remaining outlier, meaning the model makes a large prediction error for it. However, its characteristics are common (low leverage).

Influence (Cook's Distance): This is the most important finding. No observation has a Cook's distance high enough to be considered problematic. This means no single point is distorting the model's results.

#### 04. Run the model without obs 128 and plot the estimates of the regression with and without that observation

```{r}
reg3 <- lm(log(wage) ~ educ + I(educ^2) + exper + I(exper^2) + tenure + 
             I(tenure^2) + female + married + numdep + northcen + south + west, data=WAGE1 %>%
             slice(-128))

modelplot(list(reg2, reg3),
coef_omit = 'Interc')+
geom_vline(xintercept = 0,size=0.2,colour="red")
```

#### 05. Calculate the Dfbetas and plot the ones for the variable married (be careful to calculate them for the model that includes ALL observations

```{r}
DFBETAS <- dfbetas(reg2)

plot(DFBETAS[, "married"], ylab="married")
abline(h = c(-0.2, 0.2), lty = 2)
```

#### 06. Let’s check whether residuals are normal (stick with model that contains all observations)

```{r}
aaron <- reg2$residuals

kde <- kdensity(aaron, start = "normal", kernel = "gaussian")
plot(kde, main = "Kernel Density of Residuals vs. Normal Curve")
lines(kde, plot_start = TRUE, col = "red")

# Obtain a Q-Q plot
qqnorm(aaron, main = "Normal Q-Q Plot of Residuals")
qqline(aaron, col = "red", lwd = 2)

# Kolmogorov-Smirnov test
ks_test <- ks.test(aaron, "pnorm", mean = mean(aaron), sd = sd(aaron))
print(ks_test)

# Jarque-Bera test
jarque_bera_test <- jarque.bera.test(aaron)
print(jarque_bera_test)
```

The evidence from that test is more compelling, and the most reliable conclusion is still that your residuals are not normally distributed.

### Part 02: The next set of questions (7 – 9) will focus on assessing whether the chosen functional form is correct.

#### 07. Run a regression that controls for education, tenure, experience, female, married, numdep and residence dummies (basically no squared term). Then perform the RESET test using the commands:

```{r}
reg4 <- lm(log(wage) ~ educ + tenure + exper + female + married + 
                 numdep + south + northcen + west, 
                 data = WAGE1)

summary(reg4)

#Perform the RESET test including only the squared fitted term (ŷ²)
resettest(reg4, power = 2:2)

#Perform the RESET test including squared and cubed terms (ŷ² and ŷ³)
resettest(reg4, power = 2:3)

```

#### 08. Let’s go back to the specification where we control for education/tenure/experience square, and let’s test for a structural break between male and female workers. Use the dummy variable method and run two regressions.

```{r}
reg_restricted <- lm(log(wage) ~ educ + I(educ^2) + tenure + I(tenure^2) + exper + I(exper^2) +
                 married + numdep + south + northcen + west, 
                 data = WAGE1)

reg_unrestricted <- lm(log(wage) ~ female * (educ + I(educ^2) + tenure + I(tenure^2) + 
                                             exper + I(exper^2) + married + 
                                             numdep + south + northcen + west), 
                       data = WAGE1)

modelsummary(list(reg_restricted, reg_unrestricted))

#Perform the F-test to compare the models
anova(reg_restricted, reg_unrestricted)
```

#### 09. Now let’s test a linear vs a log-linear model (include the squared terms of education/tenure/experience and reintroduce the female dummy in level)

```{r}
reg5 <- lm(wage ~ educ + I(educ^2) + exper + I(exper^2) + 
                   tenure + I(tenure^2) + female + married + numdep + 
                   south + northcen + west, 
                   data = WAGE1)

reg6 <- lm(log(wage) ~ educ + I(educ^2) + exper + I(exper^2) + 
                       tenure + I(tenure^2) + female + married + numdep + 
                       south + northcen + west, 
                       data = WAGE1)

petest(reg5, reg6)
```

The PE test decisively rejects the linear model, as its p-value is nearly zero.
The test also rejects the log-linear model, but only marginally (p-value of 0.043). When both models are rejected, we choose the one that is "less wrong." Since the evidence against the linear model is far stronger, the log-linear specification is the clear winner. This choice is strongly supported by economic theory, which favors a log-linear form for wage equations.
