---
title: "Tutorial 7"
author: "Stefanio Marques - 25250375"
date: "`r Sys.Date()`"
format: pdf
editor: visual
---

### Introduction

```{r}
setwd("/Users/diegomandu/Downloads/class/programming with R/econometrics labs/tutorial 7")
```

#### Load the packages

```{r}
#| label: setup
#| message: false
#| warning: false


library(readxl)
library(tidyverse)
library(modelsummary)
library(stargazer)
library(ggplot2)
library(ggfortify)
library(car)
library(lmtest)
library(sandwich)
library(estimatr)
library(fixest)
library(ivreg)
library(gmm)
library(flextable)
```

#### 

Upload the dataset

```{r}
# or provide the full file path.
brumm <- read_excel("brumm.xlsx")


# H0: β1=0, β2=1, β3=-1
Hnull_strong <- c("(Intercept)=0", "money=1", "output=-1")
```

where INFLAT is the growth rate of the general price level, MONEY is the growth rate of
the money supply, and OUTPUT is the growth rate of national output.
The dataset you have available consists of 1995 data on 76 countrie

### Part A - OLS Estimation

```{r}
# 1. Run the OLS model [cite: 18]
ols <- lm(inflat ~ money + output, data = brumm)

# 2. Display the OLS results
summary(ols)
```

### Part B - Test Quantity Theory (OLS)
We perform the Wald test for the "strong" hypothesis  using the standard OLS standard errors.

```{r}
# Get the chi-square statistic (Wald test)
linearHypothesis(ols, Hnull_strong)

# Weak
Hnull_w <- c("money=1", "output=-1")
linearHypothesis(ols, Hnull_w)
```

### Part C - Heteroskedasticity Tests
We examine the residuals visually and with a formal test.

```{r}
# 1. Visually examine residuals [cite: 27, 28]
autoplot(ols) +
  theme_bw(base_size = 14)

# 2. Perform Breusch-Pagan test [cite: 30]
bptest(ols)
```

Conclusion: If the bptest p-value is less than 0.05, we reject the null of homoskedasticity and conclude that heteroskedasticity is present.

### Part D - Test Quantity Theory (Robust OLS)

We test the hypothesis again, this time using robust standard errors (HC3) to account for potential heteroskedasticity.

```{r}
# Test using robust standard errors [cite: 34]
linearHypothesis(ols, Hnull_strong, vcov. = hccm(ols, type = "hc3"))
linearHypothesis(ols, Hnull_w, vcov. = hccm(ols, type = "hc3"))
```

### Part E - 2SLS Estimation (with ivreg)

It is argued that output may be endogenous. We use the four proposed instruments to run a 2SLS model using the ivreg package .

The ivreg syntax is: y ~ exo_vars + endo_vars | exo_vars + instruments

```{r}
# 1. Obtain 2SLS estimator
iv <- ivreg(inflat ~ money + output | money + initial + school + inv + poprate, 
                  data = brumm)
summary(iv, diagnostics = TRUE)

# 2. Get 2SLS estimates with robust standard errors
coeftest(iv, vcov = vcovHC(iv, type = "HC3"))
```

### Part F - 2SLS Estimation (with fixest)

Here we estimate the same model using the fixest package, which has a different and very powerful syntax.

```{r}
# 1. Obtain 2SLS estimator using feols()
# We use '1' for the fixed_effects part to signify just an intercept.
model2 <- feols(inflat ~ money | 1 | output ~ initial + school + inv + poprate, 
                      data = brumm)

# 2. Get robust standard errors (the default in fixest)
summary(model2)
```

### Part G - Compare OLS and 2SLS Estimates

Let's compare the point estimates from OLS, ivreg, and fixest side-by-side. modelsummary works well for this.

```{r}
modelsummary(list(ols,iv), stars = TRUE)

modelsummary(list("OLS" = ols, 
                  "2SLS (ivreg)" = iv, 
                  "2SLS (fixest)" = model2),
             stars = TRUE)
```

### Part H - Test Quantity Theory (Robust 2SLS)

Now we test the hypothesis again using our 2SLS model with robust standard errors. We can use the iv_model object.

```{r}
# Test hypothesis using the 2SLS model and robust errors
# Note: By default, linearHypothesis uses a chi-square (Wald) test.
# We can be explicit by adding test = "F" as suggested
linearHypothesis(iv, Hnull_strong, 
                 vcov. = vcovHC(iv, type = "HC3"))

linearHypothesis(iv, Hnull_w, 
                 vcov. = vcovHC(iv, type = "HC3"))
```


### Part I - 2SLS Estimation (with fixest)

Obtain the robust standard errors. You need to install estimatr and car

```{r}
iv2 <- feols(inflat ~ money | 1 | output ~ initial + school + inv + poprate, 
                      data = brumm)

# 2. Get robust standard errors (the default in fixest)
modelsummary(list(ols,iv,iv2), stars = TRUE)
```

### Part J - GMM Estimation

We can also obtain GIVE (GMM) estimates using the gmm package.


Note: The code in the worksheet  refers to a different dataset (mroz2) and is incorrectly formatted. The code below is the correct implementation for our model.

Obtain the bootstrapped s.e. Here you need the package sandwich

```{r}
mm <- cbind(brumm$money, brumm$initial, brumm$school, brumm$inv, brumm$poprate)
gmm <- gmm(inflat ~ money + output,
           x = mm, data = brumm)


modelsummary(list(ols,iv,iv2,gmm), stars = TRUE)
```

### Part K - Perform the Wu-Hauman test for endogeneity

```{r}
step1 <- lm(output ~ money + initial + school + inv + poprate,
            data = brumm)
v <- step1$res
wu_test <- lm_robust(inflat ~ money + output + v, data = brumm)
summary(wu_test)
```

### Part L - Sargan test

```{r}
res_iv <- iv$residuals
sargan <- lm(res_iv ~ money + initial + school + inv + poprate,
             data = brumm)
summary(sargan)
```

Not rejecting

### Part M - Weak instruments test

```{r}
strong_robust <- lm_robust(output ~ money + initial + school + inv + poprate,
             data = brumm)
summary(strong_robust)
Hnull <- c("initial=0", "school=0", "inv=0", "poprate=0")
linearHypothesis(strong_robust, Hnull, test="F")
```

